---
title: "p8105_hw2_th2723"
output: html_document
---
Initial setup:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### problem 1
Basic tidy work:
```{r}
Mr.TW = 
  readxl::read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster)%>%
  mutate(
    sports_balls = as.integer(sports_balls)
  )

```
Read and clean data for 2017 and 2018:
2018:
```{r}
Mr.TW_years18 = 
  readxl::read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, range = "A2:B14") %>%
  janitor::clean_names()%>%
  drop_na(total)%>%
   mutate(year = "2018"
  )
```
2017:
```{r}
Mr.TW_years17 = 
  readxl::read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names()%>%
  drop_na(total)%>%
   mutate(year = "2017"
  )
```
Now combine 2017 and 2018, and convert month to a character variable:
```{r}
Mr.TW_years = rbind(Mr.TW_years17,Mr.TW_years18)
Mr.TW_years %>%
mutate(
       month = month.name[as.numeric(month)]
 )
select(Mr.TW_years, year, month, total)
```
~Total precipitation in 2018 is `r sum(pull(Mr.TW_years18,total))`
~The number of observation in 2017 is 12 and the number of observation is als 12.
~The median number of ports balls in a dumpster in 2017 is `r median(pull(Mr.TW, sports_balls))`
~some key variables includes year, month and total

###problem 2
```{r}
pols = 
  read_csv("fivethirtyeight_datasets/pols-month.csv")%>%
  janitor::clean_names()%>%
  separate(col= mon, sep="-", into = c("year", "month", "day"))%>%
  mutate(
         president = ifelse(prez_gop !=0, "gop","dem"),
         month = month.name[as.numeric(month)],
         year = as.integer(year),
         day = as.integer(day)
  )
pols_tidy_data = select(pols, year, month, everything(), -prez_dem, -prez_gop, -day)
  
```
snp data:
```{r}
snp = 
  read_csv("fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names()%>%
  separate(col= date, sep="/", into = c("day", "month", "year")) %>%
  mutate(
     month = month.name[as.numeric(month)],
     year = as.integer(year),
     day = as.integer(day)
  )
snp_tidy_data = select(snp, year, month, close, -day)

```
unemployment data:
```{r}
Um = 
  read_csv("fivethirtyeight_datasets/unemployment.csv")%>%
  janitor::clean_names() %>%
  rename(January = "jan", Feburary = "feb", March = "mar", April = "apr", May = "may", June = "jun", July="jul", August = "aug", September = "sep", October = "oct", November = "nov", December ="dec" )
Um_tidy_data = 
    pivot_longer(
        Um, 
        January:December,
        names_to = "month", 
        values_to = "Unemployment Rate") %>%
  mutate(
    year = as.integer(year)
  )
  
```

Merge datasets:
merge snp into pols:
```{r}
p_s = left_join(pols_tidy_data,snp_tidy_data,by = c("year", "month"))
```
merge unemployment into p_s:
```{r}
all = left_join(p_s, Um_tidy_data,by = c("year", "month"))
```
In this question, I cleared data for pols, snp and unemployment rate.
For pols, its clean data has 822 observation and 9 variables. The key variable "year" ranges from 1947 to 2015. Some key variables include president, a new variable we create for this homework. The untidy data use 1 or 0 to predict republican-dominance and democrat-dominance. However, I changed it to "dem" to indicate its dominance.
For snp, it has 787 observations and 3 variables. The key varaibles "year" ranges from 1950 to 2015. I found that "close" variable has an increasing trending over the years which indicates a general economic growth.
For unemployment rate, it has 816 observations and 3 variables. They key variable "year" ranges from 1948 to 2015. I discovered that the unemployment rate has a trend of decreasing since the Great recession, which means that the economy in United States is recovering.
### problem 3
Basic data cleaning:
```{r}
Names = 
  read_csv("Popular_Baby_Names.csv")%>%
  janitor::clean_names()%>%
  distinct()%>%
  mutate(
  ethnicity = recode(ethnicity, 'ASIAN AND PACI' = "ASIAN AND PACIFIC ISLANDER", 'WHITE NON HISP'= "WHITE NON HISPANIC", 'BLACK NON HISP' = "BLACK NON HISPANIC"  )  #Use unique(pull(Names,ethnicity)) to figure out how many expressions of the same ethnicity
  ) %>%
  mutate(
    childs_first_name = str_to_title(childs_first_name)
  )
  
```
rank in popularity of the name "Olivia"
```{r}
Olivia = 
  filter(Names, childs_first_name == "Olivia") %>%
  select(year_of_birth, ethnicity,rank) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  )
Olivia
```
Most popular name among male children:
```{r}

Male = 
  filter(Names, gender == "MALE", rank == "1") %>%
  select(year_of_birth,ethnicity,childs_first_name)%>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name
  )
```

scatter plot:
```{r}
Names %>% 
  filter(gender == "MALE") %>%
  filter(ethnicity == "WHITE NON HISPANIC")%>%
  filter(year_of_birth == "2016")%>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point(aes(), alpha = .5) +
labs(
    title = "Names for White Non-hispanic Children in 2016",
    x = "rank of name",
    y = "count of name"
)
```






